The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-07 Fsing/FThresholdsPaper/FThresholds.tex
Updates to SNC subsection of paper

\subsection*{The function {\tt isSimpleNormalCrossings}}

We briefly describe the algorithm that checks whether the polynomial $f$ factors as a product of functions $f = \prod_i f_i^{a_i}$ that are in simple normal crossing with each other, in which case, the $F$-pure threshold is just the minimum of the $1/a_i$. Recall that a set of elements $\{f_i\}$ is said to be in simple normal crossing at a point, if locally, the $f_i$ form part of a regular system of parameters.  The functionality verifies this by computing the Jacobian matrix of each subset of the $f_i$ (evaluated at the origin), and checking that these matrices have the expected rank.  It is also verified that these subsets generate ideals of the correct height.  Note that {\tt factor f} is called, which may be slow.  The user can avoid this factorization call by setting {\tt UseSpecialAlgorithms} to {\tt false}.

This method for determining whether $f$ is a product of factors that are in simple normal crossing is implemented in the function {\tt isSimpleNormalCrossing}, which is also exposed to the user.
Note that if one sets the option {\tt IsLocal} for this function to {\tt true}, then {\tt isSimpleNormalCrossing} checks that the $f_i$ are in simple normal crossing \emph{everywhere}, which can be much slower (as Jacobian ideals are computed).

>>>>

\subsection*{The function {\tt isSimpleNormalCrossings}}
As mentioned earlier, 
when the option {\tt useSpecialFunctions} for the function {\tt fpt} is set to {\tt true}, if the polynomial $f$ is neither a diagonal polynomial, a binomial, or a form in two variables, there is a final check to verify whether $f$ is a product of factors in simple normal crossing.

Suppose that $f$ has factorization $f_1^{a_i} f_2^{a_2} \cdots f_n^{a_n}$.  Recall that its factors $f_i$ are said to be in 
\emph{simple normal crossing} at a point if, locally, they form part of a regular system of parameters.  The function {\tt isSimpleNormalCrossings} determines whether $f$ has simple normal crossings by computing the Jacobian matrix of each subset of $\{ f_i \}_{i=1}^n$ (evaluated at the origin), and checking that these matrices have the expected rank, and that these subsets generate ideals of the appropriate height.

i1 : R = ZZ/7[x, y, z];

i2 : isSimpleNormalCrossing(x^3*y^2)

o2 = true

i3 : isSimpleNormalCrossing(x^2-y^2)

i4 : isSimpleNormalCrossing(x*y*(x-y))

o4 = false

i5 : isSimpleNormalCrossing(x^2-y*z)

o5 = false
\end{MyVerbatim}
}
\medspace

%             R = ZZ/7[x,y,z];
%             isSimpleNormalCrossing(x^3*y^2)
%             isSimpleNormalCrossing(x^2-y^2)
%             isSimpleNormalCrossing(x*y*(x-y))
%             isSimpleNormalCrossing(x^2-y*z)

If {\tt isSimpleNormalCrossing(f)} outputs true, then the $F$-pure threshold is easily computed, since is just the minimum of the values $1/a_i$, for $1 \leq i \leq n$.
Note that the function {\tt factor} is called whenever {\tt isSimpleNormalCrossing} is used, which can sometimes make the verification slow.  The user can avoid this when using {\tt fpt} by setting {\tt UseSpecialAlgorithms} to {\tt false}.

The function {\tt isSimpleNormalCrossing} is exposed to the user, so can be used independent of any $F$-pure threshold calculation.  
If the user sets its option {\tt IsLocal} to {\tt true} (its default value is {\tt false}), then the function checks whether the $f_i$ are in simple normal crossing \emph{everywhere}, which can be much slower since Jacobian ideals are computed. Consider the following example illustrating this functionality.

i1 : R = QQ[x, y, z];

i2 : f = (y - (x-1)^2)*y^2; --SNC at the origin, but not globally

i3 : isSimpleNormalCrossing(f)

i4 : isSimpleNormalCrossing(f, IsLocal=>false)

o4 = false

i5 : g = (y-1)^2+(x-7)^2 --doesn't even pass through the origin

       2    2
o5 = x  + y  - 14x - 2y + 50

o5 : R

i6 : isSimpleNormalCrossing(g)

o6 = true

i7 : isSimpleNormalCrossing(g, IsLocal=>false)

o7 = false

i8 : h = x*y^2*(x+1)^3*(y-1)^4; --SNC everywhere

i9 : isSimpleNormalCrossing(h)

o9 = true

i10 : isSimpleNormalCrossing(h, IsLocal=>false)

%             R = QQ[x,y,z];

%             f = (y - (x-1)^2)*y^2; --SNC at the origin, but not globally

%             isSimpleNormalCrossing(f)

%             isSimpleNormalCrossing(f, IsLocal=>false)

%             g = (y-1)^2+(x-7)^2 --doesn't even pass through the origin

%             isSimpleNormalCrossing(g)

%             isSimpleNormalCrossing(g, IsLocal=>false)

%             h = x*y^2*(x+1)^3*(y-1)^4; --SNC everywhere

%             isSimpleNormalCrossing(h)

%             isSimpleNormalCrossing(h, IsLocal=>false)
